---
title: "Basic Visualizations"
author: "Victoria DelaCruz"
date: ""
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Welcome to Basic Visualizations! We're going to go over some code on how to create a scatter plot, bar chart, and a histogram using our SDR 2023 data. 

# Starter Information

You should have *R, R Studio, ...*
**NA VALUES**

## Loading Libraries

Let's load in the `tidyverse`, `janitor`, and `here` libraries as we'll be using different functions within these packages. 

```{r}
library(tidyverse)
library(here)
library(janitor)
```

## Reading in Data

Next, let's read in our data and assign it to the variable `sdr_data`.

```{r}
sdr_data <- read.csv(here("data/SDR-2023-Data.csv"))
```

Use the `clean_names` function to clean up the column names in our data. 

```{r}
sdr_data <- sdr_data %>% 
  clean_names()
```


# Scatterplot

In the following code, we're choosing SDG Goal 3 as our x-axis and SDG Goal 4 as our y-axis. We're interested in looking at patterns within regions, so we're going to color the points by region.

```{r}
ggplot(sdr_data, aes(x= goal_3_score, y = goal_4_score, col= regions_used_for_the_sdr)) +
  geom_point() +
  labs(title = "Scatterplot", # customizing our scatterplot
       x = "SDG 3 Score",
       y = "SDG 4 Score",
       color = "Region") +
  theme_minimal() # selecting a color palette
```

Interpretation:
Many countries in Sub-Saharan Africa seem to be in the lower left quadrant of our graph, meaning they aren't doing as well in good health and well-being (SDG 3) and quality education (SDG 4) compared to the rest of the world. There seems to be a cluster or group of countries that belong to the OECD region on the top right quadrant, meaning they're doing better in terms of these goals. 

# Bar Chart
```{r}
ggplot(sdr_data, aes(x= goal_3_score, y = country)) +
  geom_bar(stat = "identity")
```
```{r}
OECD <- sdr_data %>% filter(regions_used_for_the_sdr == "OECD")

ggplot(OECD, aes(x= goal_3_score, y = country)) +
  geom_bar(stat = "identity") 
```

## Faceted Bar Chart
```{r}
unique_colors <- c("green" = "darkseagreen", "orange" = "darkorange1", "red" = "coral2", "yellow" = "darkgoldenrod2")

goal_3_faceted_bar_plot <- ggplot(data = sdr_data, aes(x = goal_3_score, 
                            y = reorder(country, goal_3_score),
                            fill = goal_3_dash)) +
  geom_col(stat = "identity") +
  facet_wrap(~regions_used_for_the_sdr, scales = "free_y") +
  scale_fill_manual(values = unique_colors) + 
  theme_minimal() +
  theme(axis.text.y = element_text(size = 4)) +
  labs(x = "SDG 3 Score",
       y = "")

goal_3_faceted_bar_plot
```

Interpretation:
As we can see in the Oceania region, there is not much data. Sub-Saharan Africa appears to have the most "red" compared to the other regions.

# Histogram
```{r}
ggplot(sdr_data, aes(x = goal_4_score)) +
  geom_histogram()
```

```{r}
ggplot(sdr_data, aes(x = goal_4_score, fill=regions_used_for_the_sdr)) +
  geom_histogram()
```

Interpretation:
Countries that fall in the higher end of the Goal 4 spectrum belong to the OECD region, while countries in Sub-Saharan Africa lie towards the lower end. 


